---
title: "ROAR-data f_p Analysis Short"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

# Load Packages

Let's first load the packages that we need for this chapter.

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
library(dplyr)
library('ggpubr')
library(ggplot2)
library(caTools)
library(ROCR) 
library(caret)
source("../code/preprocessData_new_PA.R")
source("../code/preprocessData_new_SWR.R")
library("modelr")
library(pROC)
library(viridis)
library(gridExtra)
library(sjPlot)
library(mgcv)
```

# Settings

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")
# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

# Data

## Data Loading

Load f_p data

```{r}
df.fp_data <- read.csv("~/Documents/Data/school_deidentified_data/school_ca_01/fp_data.csv")
```

Load roar data

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# all grade data

df.trials_swr <- read_csv("~/Documents/Data/school_deidentified_data/school_ca_01/Fall2022/roar_swr_trials.csv") 
df.trials_pa <- read_csv("~/Documents/Data/school_deidentified_data/school_ca_01/Fall2022/roar_pa_trials.csv")
df.trials_pa_more <- read_csv("~/Documents/Data/school_deidentified_data/school_ca_01/Fall2022/roar_pa_trials_add.csv") 
```

## Data Wrangling

```{r}
df.distinct_trials_pa <- dropDuplicate(df.trials_pa %>% filter(completed == TRUE))
df.score_pa <- generateScoreFile(df.distinct_trials_pa)
df.score_details_pa <- generateDetailedScoreFile(df.distinct_trials_pa)

df.clean_pa_firebase <- df.score_details_pa %>% 
  rename(pa_score = "roar_score") %>% 
  select(subj, FSM_correct, LSM_correct, DEL_correct, pa_score)

df.score_pa <- df.trials_pa_more %>% 
  select(-c(...1, ...2, timestmp, session)) %>% 
  pivot_longer(cols = contains("correct")) %>% 
  mutate(name = substring(name, 9,11)) %>% 
  group_by(subj, name) %>% 
  summarise(score = sum(value)) %>% 
  pivot_wider(names_from = "name", values_from = "score") %>% 
  rename(DEL_correct = "DEL", LSM_correct = "LSM", FSM_correct = "FSM") %>% 
  mutate(pa_score = DEL_correct + FSM_correct + LSM_correct ) %>% 
  filter(! is.na(pa_score)) %>% 
  select(subj, FSM_correct, LSM_correct, DEL_correct, pa_score) %>% 
  rbind(df.clean_pa_firebase) %>% 
  arrange(subj) %>% 
  rename(pid = 'subj')
  
df.distinct_trials_swr <- dropDuplicateSWR(df.trials_swr)
df.score_swr <- getSWRthetaEstimate(df.distinct_trials_swr) %>% 
  filter(trialNumTotal == 100)

length(df.score_swr$pid)
length(df.score_pa$pid)

df.roar_score <- df.score_swr %>%  
  select(pid, thetaEstimate, thetaSE) %>% 
  inner_join(df.score_pa , by = "pid")

```

There are 688 students who completed ROAR-SWR, and 657 students who completed ROAR-PA. And 632 students completed both tasks.

```{r}
df.combined <- df.roar_score %>% 
  left_join(df.fp_data %>% select(pid, novfpreadinglevel
, gradelevel), by = "pid") %>% 
  filter(!is.na(novfpreadinglevel))

length(df.combined$pid)
```

And 452 of them have novfpreadinglevel.

```{r}
df.data <- df.combined %>% 
  mutate(novfpreadinglevel = tolower(novfpreadinglevel)) %>% 
  mutate(new_score = if_else(novfpreadinglevel == "below a", 'a', novfpreadinglevel)) %>% 
  mutate(novfpreadinglevel_numeric = match(unlist(strsplit(new_score , split="")), letters)) %>% 
  mutate(risk = if_else(((novfpreadinglevel_numeric < 5) & (gradelevel == 1)) | ((novfpreadinglevel_numeric < 10) & (gradelevel == 2)) , 1, 0)) %>% 
  select(-new_score)
```

# Compare ROAR score with f_p data

```{r}
df.data_longer <- df.data %>% 
  pivot_longer(cols = c(FSM_correct, LSM_correct, DEL_correct), 
               names_to = "skills", 
               values_to = "count")
```

```{r}
df.data.1 <- df.data %>% filter(gradelevel == 1)
df.data.2 <- df.data %>% filter(gradelevel == 2)
```

```{r}
cor.test(df.data.1$pa_score, df.data.1$novfpreadinglevel_numeric,  method = "pearson")
cor.test(df.data.2$pa_score, df.data.2$novfpreadinglevel_numeric,  method = "pearson")
```

## Visualization

### ROAR-PA and ROAR-SWR plots
```{r}
x <- "hi"
xl <- expression(Speed ~ ms^-1 ~ by ~ impeller)
xl
```

```{r}
df.data$gradelevelF <- factor(x=df.data$gradelevel,levels=c(1,2),labels=c("First Grade", "Second Grade"))

g1 <-ggplot(df.data, aes(x = pa_score,
                     y = novfpreadinglevel_numeric, color= thetaEstimate)) + 
  geom_point() + scale_color_viridis(option = 'plasma')+
  xlim(0, 60) +
  labs(x = "ROAR-PA total correct",
       y = "Fountas & Pinnell") +
  scale_y_continuous(breaks = c(2,5,8,11,14),labels=c('B', 'E', 'H','K','N'))+
  geom_smooth(method = "lm",color = "black") +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..), color = "black", 
           geom = "label",vjust = .6) +
  facet_grid(cols = vars(gradelevelF))+
  labs(color = 'ROAR-SWR\ntheta')

g2 <- ggplot(df.data, mapping = aes(x = thetaEstimate,
                     y = novfpreadinglevel_numeric, color=pa_score)) + 
  geom_point() + 
  labs(x = "ROAR-SWR theta",
       y = "Fountas & Pinnell") +
  geom_smooth(method = "lm", color="black") +
  scale_color_viridis(option = 'plasma')+
  scale_y_continuous(breaks = c(2,5,8,11,14),labels=c('B', 'E', 'H','K','N'))+
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..), color = "black", 
           geom = "label", vjust = -.2) +
  facet_grid(cols = vars(gradelevelF)) +
  labs(color = 'ROAR-PA \ntotal correct')

g <- grid.arrange(g1, g2, nrow = 2)
ggsave('ROAR-PA_FandP.pdf',g, height = 6 , width = 6)
```

### ROAR-SWR

```{r}
g1 <- ggplot(df.data, mapping = aes(x = thetaEstimate,
                     y = novfpreadinglevel_numeric, color=pa_score)) + 
  geom_point() + 
  labs(x = "ROAR-SWR theta",
       y = "Fountas & Pinnell") +
  geom_smooth(method = "lm", color="black") +
  scale_color_viridis(option = 'plasma')+
  scale_y_continuous(breaks = c(2,5,8,11,14),labels=c('B', 'E', 'H','K','N'))+
  stat_cor(aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) +
  facet_grid(cols = vars(gradelevel)) +
  labs(color = 'ROAR-PA\ntheta')


```

```{r}
# Make a plot of each PA subtest agains F and P
ggplot(df.data_longer, mapping = aes(x = count,
                     y = novfpreadinglevel_numeric, 
                     group = skills,
                     color = skills)) + 
  geom_point() + 
  labs(x = "ROAR-PA correct percentile",
       y = "f_p reading level in Nov 2022 (numeric scale)") +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) +
  facet_grid(cols = vars(skills), rows = vars(gradelevel))

```

## Regression Model

### PA_numer_correct

```{r}
# fit regression model predicting F and P based on individual subtests
# First fit a model with just PA total and it is a good predictor
lm.g <- lm(formula = novfpreadinglevel_numeric ~ gradelevel,
         data = df.data) 
lm.PA.1 <- lm(formula = novfpreadinglevel_numeric ~ pa_score + gradelevel,
         data = df.data) 
lm.PA.2 <- lm(formula = novfpreadinglevel_numeric ~ pa_score + gradelevel+ DEL_correct + LSM_correct + FSM_correct,
         data = df.data) 
```

### Compare Models

lm.g = 0.36, lm.PA.1 = 0.56, lm.PA.2 = 0.56

```{r}
summary(lm.g)
summary(lm.PA.1)
summary(lm.PA.2)

# When we add in the sub-scores we can see that they contribute additional 
# information above the total score (it is significant but not much). But if we 
# Add grade subtests don't contribute more above grade and total
anova(lm.g, lm.PA.1, lm.PA.2)
```

ROAR-SWR: 0.73, ROAR-SWR + ROAR-PA: 0.76

```{r}
# Now we'll check wether PA gives more information than SWR
# Now fit model with PA and SWR
lm.SWR <- lm(formula = novfpreadinglevel_numeric ~ thetaEstimate + gradelevel,
         data = df.data)
lm.PA <- lm(formula = novfpreadinglevel_numeric ~ pa_score + gradelevel,
         data = df.data)
lm.PA.SWR <- lm(formula = novfpreadinglevel_numeric ~ pa_score + thetaEstimate + gradelevel, data = df.data)
                
summary(lm.SWR)
summary(lm.PA)
summary(lm.PA.SWR)
anova(lm.SWR, lm.PA.SWR)
```

### 1st grade only

```{r}
# Now we'll check wether PA gives more information than SWR
# Now fit model with PA and SWR
lm.SWR.1 <- lm(formula = novfpreadinglevel_numeric ~ thetaEstimate ,
         data = df.data %>% filter(gradelevel == 1))
lm.PA.SWR.1 <- lm(formula = novfpreadinglevel_numeric ~ pa_score + thetaEstimate ,
         data = df.data %>% filter(gradelevel == 1))
summary(lm.SWR.1)
summary(lm.PA.SWR.1)
anova(lm.SWR.1, lm.PA.SWR.1)
```

### 2nd grade only

```{r}
# Now we'll check wether PA gives more information than SWR
# Now fit model with PA and SWR
lm.SWR.2 <- lm(formula = novfpreadinglevel_numeric ~ thetaEstimate ,
         data = df.data %>% filter(gradelevel == 2))
lm.PA.SWR.2 <- lm(formula = novfpreadinglevel_numeric ~ pa_score + thetaEstimate ,
         data = df.data %>% filter(gradelevel == 2))
summary(lm.SWR.2)
summary(lm.PA.SWR.2)
anova(lm.SWR.2, lm.PA.SWR.2)
```

## Risk Classification

### Without Cross Validation 

```{r}
gam_model.swr.1 <- gam(risk ~ s(thetaEstimate), data = df.data %>% filter(gradelevel == 1), family = "binomial")

gam_model.pa.1 <- gam(risk ~ s(pa_score) , data = df.data %>% filter(gradelevel == 1), family = "binomial")

gam_model.swr.pa.1 <- gam(risk ~ te(thetaEstimate, pa_score), data = df.data %>% filter(gradelevel == 1), family = "binomial")
```


```{r}
generateModelPredictionResults.simple <- function (models, data) {
  list.name <- c()
  list.auc <- c()
  list.sensitivity <- c()
  list.specificity <- c()
  list.accuracy <- c()
  
  for(i in 1:length(models)){
    
    predictions <- predict(models[[i]], type = 'response')
    
    table <- tibble(pred = predictions, 
           Y = data$risk) %>% 
  mutate(pred = ifelse(pred <= 0.5, 0, 1))
    
    summary <- table %>% 
      mutate(acc = ifelse(pred == Y, 1, 0)) %>% 
      summarise(mean = mean(acc))
    

    cm <- confusionMatrix(factor(table$pred), factor(data$risk))
    overall <- cm$overall
    result <- cm$byClass
    
    
    list.name <- c(list.name, names(models)[i] )
    list.accuracy <- c(list.accuracy, round(as.numeric(overall['Accuracy']), digits = 3))
    list.sensitivity <- c(list.sensitivity, round(as.numeric(result['Sensitivity']), digits = 3))
    list.specificity <- c(list.specificity, round(as.numeric(result['Specificity']), digits = 3))
    
    r = roc(response = table$Y, predictor = predictions)
    list.auc <- c(list.auc, round(as.numeric(r$auc), digits = 3))
    
  }
  
  return (tibble(name = list.name, 
                 accuracy = list.accuracy,
         sensitivity = list.sensitivity, 
         specificity = list.specificity,
         auc = list.auc))
}

model.compare.1 <- generateModelPredictionResults.simple(
  list("gam_model.swr" = gam_model.swr.1, 
       "gam_model.pa" = gam_model.pa.1, 
       "gam_model.swr.pa" = gam_model.swr.pa.1), 
  df.data %>% filter(gradelevel == 1))
```

```{r}
gam_model.swr.2 <- gam(risk ~ s(thetaEstimate), data = df.data %>% filter(gradelevel == 2), family = "binomial")

gam_model.pa.2 <- gam(risk ~ s(pa_score), data = df.data %>% filter(gradelevel == 2), family = "binomial")

gam_model.swr.pa.2 <- gam(risk ~ te(thetaEstimate, pa_score), data = df.data %>% filter(gradelevel == 2), family = "binomial")

```

Model Comparison Results without cross validation:
```{r}
model.compare.2 <- generateModelPredictionResults.simple(
  list("gam_model.swr" = gam_model.swr.2, 
       "gam_model.pa" = gam_model.pa.2, 
       "gam_model.swr.pa" = gam_model.swr.pa.2), 
  df.data %>% filter(gradelevel == 2))

df.models <- model.compare.1 %>% 
  mutate(grade = 1) %>% 
  rbind(model.compare.2 %>% mutate(grade=2)) %>% 
  rename(model = "name") %>% 
  relocate(grade)

df.models
```
```{r}
tab_df(df.models, file="models.doc") 
```


# Spring 2022 data predicts f_p outcome

## read data

We have 130 students who are qualified in this analysis:\
1. completed both SWR and PA in spring 2022\
2. received a f_p score in November 2022\

```{r}
df.data_2022_spring <- read_csv("~/Documents/Data/school_deidentified_data/school_ca_01/roar_data_2022_spring.csv") 

df.fp_2021_2022 <- read_csv("~/Documents/Data/school_deidentified_data/school_ca_01/fp_data_2021_2022.csv") 
```

```{r}
# merge novfpreadinglevel into the dataframe
df.data.22.spring <- df.data_2022_spring %>% 
  left_join(df.fp_data, by = 'pid') %>% 
  mutate(novfpreadinglevel = tolower(novfpreadinglevel)) %>% 
  mutate(new_score = if_else(novfpreadinglevel == "below a", 'a', novfpreadinglevel)) %>% 
  mutate(novfpreadinglevel_numeric = match(unlist(strsplit(new_score , split="")), letters)) %>% 
  mutate(risk = if_else(((novfpreadinglevel_numeric < 5) & (gradelevel == 1)) | ((novfpreadinglevel_numeric < 10) & (gradelevel == 2)) , 1, 0)) %>% 
  filter(!is.na(novfpreadinglevel)) %>% 
  select(-c(new_score, ...1))

length(df.data.22.spring$pid)
```

```{r}
df.a <- df.fp_2021_2022 %>% 
  left_join(df.fp_data, by = 'pid') %>% 
  mutate(novfpreadinglevel = tolower(novfpreadinglevel)) %>% 
  mutate(new_score = if_else(novfpreadinglevel == "below a", 'a', novfpreadinglevel)) %>% 
  mutate(novfpreadinglevel_numeric = match(unlist(strsplit(new_score , split="")), letters)) %>% 
  mutate(risk = if_else(((novfpreadinglevel_numeric < 5) & (gradelevel == 1)) | ((novfpreadinglevel_numeric < 10) & (gradelevel == 2)) , 1, 0)) %>% 
  filter(!is.na(novfpreadinglevel)) %>% 
  select(-c(new_score, ...1))
```

```{r}
df.data.fp <- df.a %>% 
  filter(!is.na(fp_level_03_2022)) %>% 
  select(pid, fp_level_03_2022, risk) %>% 
  mutate(fp_level_03_2022 = tolower(fp_level_03_2022)) %>% 
   mutate(new_score = if_else(fp_level_03_2022 == "aa", 'a', fp_level_03_2022)) %>% 
  mutate(fp_level_03_2022_numeric = match(unlist(strsplit(new_score, split="")), letters))

g.fp.spring <- gam(risk ~ te(fp_level_03_2022_numeric), data = df.data.fp, family = binomial)

p.fp.spring = predict(g.fp.spring,  type = 'response')


r.g.fp.spring = roc(response = g.fp.spring$y, predictor = p.fp.spring)
print(r.g.fp.spring)

df.data.fp <- df.a %>% 
  filter(!is.na(fp_level_06_2022)) %>% 
  select(pid, fp_level_06_2022, risk) %>% 
  mutate(fp_level_06_2022 = tolower(fp_level_06_2022)) %>% 
   mutate(new_score = if_else(fp_level_06_2022 == "aa", 'a', fp_level_06_2022)) %>% 
  mutate(fp_level_06_2022_numeric = match(unlist(strsplit(new_score, split="")), letters))

g.fp.spring <- gam(risk ~ te(fp_level_06_2022_numeric), data = df.data.fp, family = binomial)

p.fp.spring = predict(g.fp.spring,  type = 'response')


r.g.fp.spring = roc(response = g.fp.spring$y, predictor = p.fp.spring)
print(r.g.fp.spring)
```


## Visualization

```{r}
ggplot(df.data.22.spring, mapping = aes(x = pa_score,
                     y = novfpreadinglevel_numeric)) + 
  geom_point() + 
  labs(x = "ROAR-PA total correct in Spring 2022",
       y = "f_p reading level in Nov 2022 (numeric scale)") +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) +
  facet_grid(cols = vars(gradelevel))

ggplot(df.data.22.spring, mapping = aes(x = roar_swr,
                     y = novfpreadinglevel_numeric)) + 
  geom_point() + 
  labs(x = "ROAR-SWR total correct (Block A only) in Spring 2022",
       y = "f_p reading level in Nov 2022 (numeric scale)") +
  geom_smooth(method = "lm") +
  stat_cor(aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) +
  facet_grid(cols = vars(gradelevel))
```

## GAM (without splitting)

```{r}
# classification
g.swr.spring <- gam(risk ~ te(roar_swr), data = df.data.22.spring, family = binomial)

g.swr.pa.spring <- gam(risk ~ te(roar_swr, pa_score), data = df.data.22.spring, family = binomial)

g.pa.spring <- gam(risk ~ te(pa_score), data = df.data.22.spring, family = binomial)

summary(g.swr.spring)
summary(g.pa.spring)
summary(g.swr.pa.spring)

p.swr.spring = predict(g.swr.spring,  type = 'response')
p.pa.spring = predict(g.pa.spring,  type = 'response')
p.swr.pa.spring = predict(g.swr.pa.spring,  type = 'response')

r.g.swr.spring = roc(response = g.swr.spring$y, predictor = p.swr.spring)
print(r.g.swr.spring)
r.g.pa.spring = roc(response = g.pa.spring$y, predictor = p.pa.spring)
print(r.g.pa.spring)
r.g.swr.pa.spring = roc(response = g.swr.pa.spring$y, predictor = p.swr.pa.spring)
print(r.g.swr.pa.spring)

```

GAM: use Spring 2022 roar data to predict Nov 2022 f_p\
1. AUC of ROAR_SWR: 0.8255\
2. AUC of ROAR_PA: 0.7072\
3. AUC of ROAR_SWR + ROAR_PA: 0.8385
